---
title: "QC, integration, and annotation of RVS samples"
author: "GinoBonazza (ginoandrea.bonazza@usz.ch)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r knitr config, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(dpi = 300, fig.align = "center")
```

## Setup

```{r setup, class.source = "fold-hide"}
current_file <- "RVS_QC_integration_annotation"

library(here)
library(readr)
library(readxl)
library(Seurat)
library(DropletUtils)
library(Matrix)
library(scDblFinder)
library(scCustomize)
library(dplyr)
library(ggplot2)
library(magrittr)
library(tidyverse)
library(reshape2)
library(S4Vectors)
library(SingleCellExperiment)
library(pheatmap)
library(png)
library(gridExtra)
library(knitr)
library(scales)
library(RColorBrewer)
library(Matrix.utils)
library(tibble)
library(scater)
library(patchwork)
library(statmod)
library(clustree)
library(harmony)
library(speckle)
library(qs2)

output_dir_data <- here::here("output", current_file)
if (!dir.exists(output_dir_data)) dir.create(output_dir_data)

if (!dir.exists(here::here("docs", "figure"))) dir.create(here::here("docs", "figure"))

output_dir_figs <- here::here("docs", "figure", paste0(current_file, ".Rmd"))
if (!dir.exists(output_dir_figs)) dir.create(output_dir_figs)
```

## Quality control

Load and merge CellBender and Cell Ranger output for the two RVS samples only.

```{r load data, echo=TRUE, results="hide"}
cell_bender_merged <- Read_CellBender_h5_Multi_File(
  merge = TRUE,
  data_dir = here::here("data", "Cellbender_output"),
  custom_name = "_filtered.h5",
  sample_list = c(
    "cellbender_output_RVS_200625_d0",
    "cellbender_output_RVS_200625_d14_A1"
  ),
  sample_names = c(
    "RVS_200625_d0",
    "RVS_200625_d14_A1"
  )
)

cell_ranger_merged <- Read10X_h5_GEO(
  merge = TRUE,
  data_dir = here::here("data", "Cellranger_output"),
  shared_suffix = "_filtered_feature_bc_matrix.h5",
  sample_list = c(
    "RVS_200625_d0",
    "RVS_200625_d14_A1"
  ),
  sample_names = c(
    "RVS_200625_d0",
    "RVS_200625_d14_A1"
  )
)

RVS_preQC <- Create_CellBender_Merged_Seurat(
  raw_cell_bender_matrix = cell_bender_merged,
  raw_counts_matrix = cell_ranger_merged,
  raw_assay_name = "RAW",
  cell_bender_assay_name = "RNA"
)

rm(cell_bender_merged, cell_ranger_merged)
```

Add metadata: sample, time point, mitochondrial percentage, and ribosomal percentage.

```{r add metadata}
cn <- Cells(RVS_preQC)

RVS_preQC$Sample <- dplyr::case_when(
  grepl("^RVS_200625_d0[_-]", cn) ~ "RVS_200625_d0",
  grepl("^RVS_200625_d14_A1[_-]", cn) ~ "RVS_200625_d14_A1",
  TRUE ~ NA_character_
)

RVS_preQC$time_point <- dplyr::case_when(
  RVS_preQC$Sample == "RVS_200625_d0" ~ "d0",
  RVS_preQC$Sample == "RVS_200625_d14_A1" ~ "d14",
  TRUE ~ NA_character_
)

RVS_preQC$condition <- RVS_preQC$time_point

RVS_preQC[["percent.mt"]] <- PercentageFeatureSet(RVS_preQC, pattern = "^MT-")
RVS_preQC[["percent.rp"]] <- PercentageFeatureSet(RVS_preQC, pattern = "^RP[SL]")

table(RVS_preQC$Sample, useNA = "ifany")
```

Check how much CellBender affected the dataset.

```{r Cellbender, fig.height=4, fig.width=7}
table(RVS_preQC$Sample)

RVS_preQC <- Add_CellBender_Diff(
  seurat_object = RVS_preQC,
  raw_assay_name = "RAW",
  cell_bender_assay_name = "RNA"
)

median_stats <- Median_Stats(
  seurat_object = RVS_preQC,
  group_by_var = "Sample",
  median_var = c("nCount_Diff", "nFeature_Diff")
)
median_stats

feature_diff <- CellBender_Feature_Diff(
  seurat_object = RVS_preQC,
  raw_assay = "RAW",
  cell_bender_assay = "RNA"
)
head(feature_diff)

CellBender_Diff_Plot(feature_diff_df = feature_diff)
```

Check quality control parameters before filtering.

```{r QC_pre_VlnPlots, fig.height=4, fig.width=8}
p1 <- VlnPlot(RVS_preQC, features = "nCount_RNA", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()
p2 <- VlnPlot(RVS_preQC, features = "nFeature_RNA", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()
p3 <- VlnPlot(RVS_preQC, features = "percent.mt", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()
p4 <- VlnPlot(RVS_preQC, features = "percent.rp", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

QC_pre_VlnPlots <- p1 + p2 + p3 + p4 + plot_layout(ncol = 4)
QC_pre_VlnPlots
```

```{r QC_pre_VlnPlots_zoom, fig.height=4, fig.width=8}
p1 <- VlnPlot(RVS_preQC, features = "nCount_RNA", group.by = "Sample", pt.size = 0, y.max = 20000) + theme(axis.title.x = element_blank()) + NoLegend()
p2 <- VlnPlot(RVS_preQC, features = "nFeature_RNA", group.by = "Sample", pt.size = 0, y.max = 10000) + theme(axis.title.x = element_blank()) + NoLegend()
p3 <- VlnPlot(RVS_preQC, features = "percent.mt", group.by = "Sample", pt.size = 0, y.max = 5) + theme(axis.title.x = element_blank()) + NoLegend()
p4 <- VlnPlot(RVS_preQC, features = "percent.rp", group.by = "Sample", pt.size = 0, y.max = 5) + theme(axis.title.x = element_blank()) + NoLegend()

QC_pre_VlnPlots_zoom <- p1 + p2 + p3 + p4 + plot_layout(ncol = 4)
QC_pre_VlnPlots_zoom
```

Doublet detection.

```{r doublets detection}
RVS_preQC_sce <- as.SingleCellExperiment(RVS_preQC)
RVS_preQC_sce <- scDblFinder(RVS_preQC_sce, samples = "Sample", clusters = TRUE)
table(RVS_preQC_sce@colData$scDblFinder.class)

RVS_preQC <- as.Seurat(RVS_preQC_sce, counts = "counts", data = "logcounts")
rm(RVS_preQC_sce)

RVS_preQC@meta.data[RVS_preQC@meta.data$scDblFinder.class %in% "singlet", "scDblFinder.n"] <- paste0("Singlets (n=", table(RVS_preQC$scDblFinder.class)["singlet"], ")")
RVS_preQC@meta.data[RVS_preQC@meta.data$scDblFinder.class %in% "doublet", "scDblFinder.n"] <- paste0("Doublets (n=", table(RVS_preQC$scDblFinder.class)["doublet"], ")")
RVS_preQC$scDblFinder.n <- factor(RVS_preQC$scDblFinder.n)
```

```{r QC_pre_Doublets, fig.height=4, fig.width=6}
p1 <- VlnPlot(RVS_preQC, features = "nCount_RNA", split.by = "scDblFinder.n", group.by = "Sample", pt.size = 0, y.max = 50000) + theme(axis.title.x = element_blank()) + NoLegend()
p2 <- VlnPlot(RVS_preQC, features = "nFeature_RNA", split.by = "scDblFinder.n", group.by = "Sample", pt.size = 0, y.max = 10000) + theme(axis.title.x = element_blank())

QC_pre_Doublets <- p1 + p2 + plot_layout(ncol = 2)
QC_pre_Doublets
```

```{r save RVS_preQC}
qs_save(
  RVS_preQC,
  file = here::here(output_dir_data, "RVS_preQC.qs2"),
  nthreads = 32
)
```

Remove doublets.

```{r remove doublets}
RVS <- subset(RVS_preQC, subset = scDblFinder.class == "singlet")
table(RVS$scDblFinder.class)
rm(RVS_preQC)

RVS@assays[["RAW"]] <- NULL
DefaultAssay(RVS) <- "RNA"
```

Check quality control parameters after removing doublets.

```{r QC_pre_no_doublets_VlnPlots, fig.height=4, fig.width=8}
p1 <- VlnPlot(RVS, features = "nCount_RNA", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()
p2 <- VlnPlot(RVS, features = "nFeature_RNA", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()
p3 <- VlnPlot(RVS, features = "percent.mt", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()
p4 <- VlnPlot(RVS, features = "percent.rp", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

QC_pre_no_doublets_VlnPlots <- p1 + p2 + p3 + p4 + plot_layout(ncol = 4)
QC_pre_no_doublets_VlnPlots
```

```{r QC_pre_no_doublets_VlnPlots_zoom, fig.height=4, fig.width=8}
p1 <- VlnPlot(RVS, features = "nCount_RNA", group.by = "Sample", pt.size = 0, y.max = 20000) + theme(axis.title.x = element_blank()) + NoLegend()
p2 <- VlnPlot(RVS, features = "nFeature_RNA", group.by = "Sample", pt.size = 0, y.max = 10000) + theme(axis.title.x = element_blank()) + NoLegend()
p3 <- VlnPlot(RVS, features = "percent.mt", group.by = "Sample", pt.size = 0, y.max = 5) + theme(axis.title.x = element_blank()) + NoLegend()
p4 <- VlnPlot(RVS, features = "percent.rp", group.by = "Sample", pt.size = 0, y.max = 5) + theme(axis.title.x = element_blank()) + NoLegend()

QC_pre_no_doublets_VlnPlots_zoom <- p1 + p2 + p3 + p4 + plot_layout(ncol = 4)
QC_pre_no_doublets_VlnPlots_zoom
```

Filter based on number of counts, features, and percentage of mitochondrial genes.

```{r filter}
RVS <- subset(
  RVS,
  subset = nFeature_RNA > 300 &
    percent.mt < 20 &
    nCount_RNA > 500 &
    nCount_RNA < 100000
)

table(RVS$Sample)
```

```{r cols}
cell_type_cols <- c("#d05050", "#33A02C", "#1F78B4", "#CAB2D6", "#FF7F00", "#B2DF8A", "#B15928", "#FFD000", "#FA95C0", "#FDBF6F", "#6A3D9A", "#7aCDF9", "#E6E600", "#8DD3C7", "#FB8072")
sample_cols <- c(
  "RVS_200625_d0"     = "#BDBDBD",
  "RVS_200625_d14_A1" = "#B85C4A"
)
cluster_cols <- DiscretePalette_scCustomize(num_colors = 20, palette = "polychrome", shuffle_pal = TRUE, seed = 220510)
```

```{r plot_cols, fig.height=3, fig.width=10}
PalettePlot(cell_type_cols) + PalettePlot(sample_cols) + PalettePlot(cluster_cols)
```

Check quality control parameters after filtering.

```{r QC_post_VlnPlots, fig.height=4, fig.width=8}
p1 <- VlnPlot(RVS, features = "nCount_RNA", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()
p2 <- VlnPlot(RVS, features = "nFeature_RNA", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()
p3 <- VlnPlot(RVS, features = "percent.mt", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()
p4 <- VlnPlot(RVS, features = "percent.rp", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

QC_post_VlnPlots <- p1 + p2 + p3 + p4 + plot_layout(ncol = 4)
QC_post_VlnPlots
```

```{r QC_post_VlnPlots_zoom, fig.height=4, fig.width=8}
p1 <- VlnPlot(RVS, features = "nCount_RNA", group.by = "Sample", pt.size = 0, y.max = 20000) + theme(axis.title.x = element_blank()) + NoLegend()
p2 <- VlnPlot(RVS, features = "nFeature_RNA", group.by = "Sample", pt.size = 0, y.max = 10000) + theme(axis.title.x = element_blank()) + NoLegend()
p3 <- VlnPlot(RVS, features = "percent.mt", group.by = "Sample", pt.size = 0, y.max = 5) + theme(axis.title.x = element_blank()) + NoLegend()
p4 <- VlnPlot(RVS, features = "percent.rp", group.by = "Sample", pt.size = 0, y.max = 5) + theme(axis.title.x = element_blank()) + NoLegend()

QC_post_VlnPlots_zoom <- p1 + p2 + p3 + p4 + plot_layout(ncol = 4)
QC_post_VlnPlots_zoom
```

## Integration and clustering

Normalize and scale using SCTransform.

```{r SCTransform, message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
DefaultAssay(RVS) <- "RNA"
RVS <- SplitObject(RVS, split.by = "Sample")

for (i in 1:length(RVS)) {
  RVS[[i]] <- SCTransform(
    RVS[[i]],
    vst.flavor = "v2",
    vars.to.regress = c("percent.mt"),
    return.only.var.genes = TRUE,
    variable.features.n = 3000,
    method = "glmGamPoi"
  )
}
```

PCA.

```{r PCA}
features <- SelectIntegrationFeatures(object.list = RVS, nfeatures = 3000)
RVS <- merge(x = RVS[[1]], y = RVS[2:length(RVS)], merge.data = TRUE)
VariableFeatures(RVS) <- features
RVS <- RunPCA(RVS, npcs = 50)
```

```{r Elbow_not_integrated, fig.width=4, fig.height=4}
ElbowPlot(RVS, ndims = 50)
```

Clustering without integration.

```{r clustering not integrated}
DefaultAssay(RVS) <- "SCT"
RVS <- RunUMAP(RVS, dims = 1:50)
RVS <- FindNeighbors(RVS, dims = 1:50)
RVS <- FindClusters(RVS, resolution = seq(0.1, 0.8, by = 0.1))
```

```{r Clustree_not_integrated, fig.width=6, fig.height=9}
clustree::clustree(
  RVS@meta.data[, grep("SCT_snn_res", colnames(RVS@meta.data))],
  prefix = "SCT_snn_res."
)
```

```{r UMAPs_not_integrated, fig.width=10, fig.height=4}
DimPlot(
  RVS,
  reduction = "umap",
  shuffle = TRUE,
  group.by = c("SCT_snn_res.0.3", "Sample"),
  ncol = 2,
  label = TRUE
)
```

Integrate the samples using Harmony.

```{r RunHarmony}
RVS <- RunHarmony(
  RVS,
  assay.use = "SCT",
  group.by.vars = "Sample",
  dims.use = 1:50,
  plot_convergence = TRUE,
  .options = harmony_options(max.iter.cluster = 100)
)
```

Clustering after integration.

```{r clustering integrated}
DefaultAssay(RVS) <- "SCT"
RVS <- RunUMAP(RVS, dims = 1:50, reduction = "harmony")
RVS <- FindNeighbors(RVS, dims = 1:50, reduction = "harmony")
RVS <- FindClusters(RVS, resolution = seq(0.1, 0.8, by = 0.1))
```

```{r Clustree_integrated_low_quality, fig.width=6, fig.height=9}
clustree::clustree(
  RVS@meta.data[, grep("SCT_snn_res", colnames(RVS@meta.data))],
  prefix = "SCT_snn_res."
)
```

```{r UMAPs_integrated_low_quality, fig.width=10, fig.height=4}
DimPlot(
  RVS,
  reduction = "umap",
  shuffle = TRUE,
  group.by = c("SCT_snn_res.0.7", "Sample"),
  ncol = 2,
  label = FALSE
)
```
```{r UMAPs_integrated_low_quality_label, fig.width=10, fig.height=4}
DimPlot(
  RVS,
  reduction = "umap",
  shuffle = TRUE,
  group.by = c("SCT_snn_res.0.7", "Sample"),
  ncol = 2,
  label = TRUE
)
```

Find marker genes for the integrated clusters before possible removal of low-quality cells.

```{r FindAllMarkers low quality}
DefaultAssay(RVS) <- "RNA"
RVS <- NormalizeData(RVS)
Idents(RVS) <- "SCT_snn_res.0.7"

Markers <- FindAllMarkers(RVS, only.pos = TRUE, min.pct = 0.5, logfc.threshold = 1)
write.csv(Markers, here::here(output_dir_data, "RVS_Markers_all_low_quality.csv"))

Markers_top10 <- as.data.frame(Markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC))
write.csv(Markers_top10, here::here(output_dir_data, "RVS_Markers_top10_low_quality.csv"))

Markers_top3 <- as.data.frame(Markers %>% group_by(cluster) %>% top_n(n = 3, wt = avg_log2FC))
write.csv(Markers_top3, here::here(output_dir_data, "RVS_Markers_top3_low_quality.csv"))
```

```{r Heatmap_low_quality, fig.width=14, fig.height=10}
heatmap_cols <- colorRampPalette(RColorBrewer::brewer.pal(9, "RdBu"))(256)
heatmap_cols <- rev(heatmap_cols[1:256])

Heatmap <- DoHeatmap(
  RVS,
  label = FALSE,
  draw.line = FALSE,
  features = Markers_top10$gene,
  assay = "SCT"
) +
  scale_fill_gradientn(colours = heatmap_cols) +
  theme(axis.text.y = element_text(size = 5)) +
  theme(plot.margin = unit(c(0.1, 0, 0, 0), "inches"))

Heatmap
```

Check QC parameters in each cluster.

```{r QC_VlnPlots_low_quality, fig.width=12, fig.height=6}
VlnPlot(
  RVS,
  features = c("nCount_RNA", "nFeature_RNA", "percent.mt", "percent.rp", "scDblFinder.score"),
  ncol = 3,
  pt.size = 0
) + theme(axis.title.x = element_blank())
```

Check cell cycle genes.

```{r Cell_cycle_UMAP, fig.width=5, fig.height=4}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
RVS <- CellCycleScoring(RVS, s.features = s.genes, g2m.features = g2m.genes)

DimPlot(
  RVS,
  reduction = "umap",
  shuffle = TRUE,
  group.by = "Phase"
)
```

## Cell type annotation

Placeholder annotation based on resolution 0.3. Change this after checking the marker genes.

```{r annotation}

cluster_annotation <- c( "0" = "CM I", "1" = "EC I", "2" = "EC II", "3" = "Fibroblasts", "4" = "Pericytes", "5" = "CM high MT I", "6" = "CM high MT II", "7" = "CM high MT III", "8" = "Macrophages", "9" = "Smooth Muscle Cells", "10" = "T cells", "11" = "CM II", "12" = "EC III", "13" = "EC IV", "14" = "Neuronal" ) 
RVS$cell_type <- plyr::mapvalues( x = as.character(RVS$SCT_snn_res.0.7), from = names(cluster_annotation), to = unname(cluster_annotation) )

cell_type_counts <- sort(table(RVS$cell_type), decreasing = TRUE)

cm_levels <- c(
  "CM I",
  "CM II",
  "CM high MT I",
  "CM high MT II",
  "CM high MT III"
)

ec_levels <- c(
  "EC I",
  "EC II",
  "EC III",
  "EC IV"
)

other_levels <- names(cell_type_counts)[
  !names(cell_type_counts) %in% c(cm_levels, ec_levels)
]

new_cell_type_levels <- c(
  cm_levels,
  ec_levels,
  other_levels
)

RVS$cell_type <- factor(
  RVS$cell_type,
  levels = new_cell_type_levels
)

Idents(RVS) <- RVS$cell_type

table(RVS$cell_type, useNA = "ifany")
```

```{r}
RVS <- qs_read(file = here::here(output_dir_data, "RVS.qs2"),
  nthreads = 32)
```

```{r QC_sample, fig.width=12, fig.height=8}
VlnPlot_scCustom(
  RVS,
  features = c("nCount_RNA", "nFeature_RNA", "percent.mt", "percent.rp", "scDblFinder.score"),
  group.by = "cell_type",
  num_columns = 3,
  pt.size = 0,
  colors_use = cell_type_cols
) & theme(axis.title.x = element_blank())
```

```{r UMAP_cluster, fig.height=5, fig.width=5}
DimPlot_scCustom(RVS, group.by = "SCT_snn_res.0.3", colors_use = cluster_cols, label = TRUE)
```

```{r UMAP_sample, fig.width=5, fig.height=5}
DimPlot(
  RVS,
  group.by = "Sample",
  reduction = "umap",
  cols = sample_cols,
  label = FALSE,
  shuffle = TRUE,
  pt.size = 1
) +
  NoLegend() +
  theme(
    axis.text = element_text(size = 8, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    plot.title = element_blank()
  )
```

```{r UMAP_sample_split, fig.width=8, fig.height=4}
DimPlot(
  RVS,
  group.by = "Sample",
  split.by = "Sample",
  reduction = "umap",
  cols = sample_cols,
  label = FALSE,
  shuffle = TRUE,
  pt.size = 0.5
) +
  NoLegend() +
  theme(
    axis.text = element_text(size = 8, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    strip.text = element_text(size = 11, face = "bold")
  )
```

```{r UMAP_cell_type, fig.width=5, fig.height=5}
p <- DimPlot(
  RVS,
  group.by = "cell_type",
  reduction = "umap",
  cols = cell_type_cols,
  label = FALSE,
  shuffle = TRUE,
  pt.size = 1
) +
  NoLegend() +
  theme(
    axis.text = element_text(size = 8, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    plot.title = element_blank()
  )

LabelClusters(
  p,
  id = "cell_type",
  size = 4,
  repel = TRUE
)
```

```{r UMAP_percent_mt, fig.width=5, fig.height=5}
FeaturePlot(
  RVS,
  features = "percent.mt",
  reduction = "umap",
  pt.size = 1,
  order = TRUE
) +
  scale_color_gradientn(
    colors = c("grey90", "#FDBF6F", "#d05050")
  ) +
  NoLegend() +
  theme(
    axis.text = element_text(size = 8, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    plot.title = element_blank()
  )
```

```{r}
get_unique_top_markers <- function(markers, n = 10, cluster_col = "cluster", gene_col = "gene") {
  
  markers <- markers %>%
    dplyr::arrange(.data[[cluster_col]], dplyr::desc(avg_log2FC))
  
  clusters <- unique(markers[[cluster_col]])
  used_genes <- character(0)
  out <- list()
  
  for (cl in clusters) {
    
    cl_markers <- markers %>%
      dplyr::filter(.data[[cluster_col]] == cl) %>%
      dplyr::filter(!.data[[gene_col]] %in% used_genes) %>%
      dplyr::slice_head(n = n)
    
    used_genes <- c(used_genes, cl_markers[[gene_col]])
    out[[as.character(cl)]] <- cl_markers
  }
  
  dplyr::bind_rows(out)
}
```

```{r FindAllMarkers cell_type}
DefaultAssay(RVS) <- "RNA"
Idents(RVS) <- "cell_type"

Markers <- FindAllMarkers(RVS, only.pos = TRUE, min.pct = 0.5, logfc.threshold = 1)
write.csv(Markers, here::here(output_dir_data, "RVS_Markers_all_cell_type.csv"))

Markers_top10 <- get_unique_top_markers(
  markers = Markers,
  n = 10,
  cluster_col = "cluster",
  gene_col = "gene"
)
write.csv(Markers_top10, here::here(output_dir_data, "RVS_Markers_top10_cell_type.csv"))

Markers_top3 <- get_unique_top_markers(
  markers = Markers,
  n = 3,
  cluster_col = "cluster",
  gene_col = "gene"
)
write.csv(Markers_top3, here::here(output_dir_data, "RVS_Markers_top3_cell_type.csv"))
```

```{r Heatmap_cell_type, fig.width=12, fig.height=10}
genes.use <- intersect(Markers_top10$gene, rownames(RVS[["SCT"]]@scale.data))
DefaultAssay(RVS) <- "SCT"

Heatmap <- DoHeatmap(
  RVS,
  features = genes.use,
  label = FALSE,
  draw.line = TRUE,
  group.by = "cell_type",
  group.colors = cell_type_cols,
  lines.width = 1
) +
  scale_fill_gradientn(colours = heatmap_cols) +
  theme(text = element_text(size = 15), axis.text.y = element_text(size = 6))

Heatmap
```

```{r DotPlot_markers, fig.width=12, fig.height=6}
Idents(RVS) <- factor(Idents(RVS), levels = rev(levels(Idents(RVS))))

p <- DotPlot_scCustom(
  RVS,
  features = unique(Markers_top3$gene),
  x_lab_rotate = TRUE
)

p + theme(
  axis.text.x = element_text(size = 12, angle = 60, vjust = 1, hjust = 1),
  axis.text.y = element_text(size = 15)
)

Idents(RVS) <- RVS$cell_type
```

```{r FeatPlot_canonical_markers, fig.width=14, fig.height=9}
FeaturePlot_scCustom(
  RVS,
  features = c(
    "RYR2", "TTN", "COL1A1", "DCN", "VWF", "PECAM1", "RGS5",
    "MRC1", "CD68", "PTPRC", "RGS6"
  ),
  num_columns = 4,
  order = FALSE
)
```

```{r Barplot_Sample, fig.height=7, fig.width=8, message=FALSE}
props <- getTransformedProps(RVS$cell_type, RVS$Sample, transform = "logit")

par(mar = c(14, 5, 1, 18), xpd = TRUE)
barplot(
  props$Proportions,
  legend = FALSE,
  ylab = "Proportions",
  col = cell_type_cols,
  cex.names = 1.4,
  las = 2,
  font.lab = 2,
  font.axis = 2,
  cex.axis = 1,
  cex.lab = 1.4
)
legend(
  "topright",
  inset = c(-0.8, -0),
  legend = rownames(props$Proportions),
  fill = cell_type_cols,
  bty = "n",
  cex = 1.2
)
```

```{r Barplot_cell_type, fig.height=5, fig.width=10, message=FALSE}
props <- getTransformedProps(RVS$Sample, RVS$cell_type, transform = "logit")

par(mar = c(12, 5, 1, 14), xpd = TRUE)
barplot(
  props$Proportions,
  legend = FALSE,
  ylab = "Proportions",
  col = sample_cols,
  cex.names = 1.2,
  las = 2,
  font.lab = 2,
  font.axis = 2,
  cex.axis = 1,
  cex.lab = 1.4
)
legend(
  "topright",
  inset = c(-0.4, 0.3),
  legend = rownames(props$Proportions),
  fill = sample_cols,
  bty = "n",
  cex = 1.2
)
```

```{r summary tables}
dim(RVS)
table(RVS$Sample)
table(RVS$cell_type)
table(RVS$Sample, RVS$cell_type)
```

```{r cell_type_count_heatmap_table, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=3}
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)

cell_type_levels <- levels(RVS$cell_type)

cell_type_count_long <- RVS@meta.data %>%
  dplyr::count(Sample, cell_type, name = "Nuclei") %>%
  tidyr::complete(
    Sample,
    cell_type = cell_type_levels,
    fill = list(Nuclei = 0)
  ) %>%
  dplyr::mutate(
    Sample = dplyr::case_when(
      Sample == "RVS_200625_d0" ~ "d0",
      Sample == "RVS_200625_d14_A1" ~ "d14",
      TRUE ~ as.character(Sample)
    )
  )

# add total column
sample_totals <- cell_type_count_long %>%
  dplyr::group_by(Sample) %>%
  dplyr::summarise(Nuclei = sum(Nuclei), .groups = "drop") %>%
  dplyr::mutate(cell_type = "Total")

# add total row
cell_type_totals <- cell_type_count_long %>%
  dplyr::group_by(cell_type) %>%
  dplyr::summarise(Nuclei = sum(Nuclei), .groups = "drop") %>%
  dplyr::mutate(Sample = "Total")

# add grand total
grand_total <- tibble(
  Sample = "Total",
  cell_type = "Total",
  Nuclei = sum(cell_type_count_long$Nuclei)
)

plot_df <- dplyr::bind_rows(
  cell_type_count_long,
  sample_totals,
  cell_type_totals,
  grand_total
) %>%
  dplyr::mutate(
    cell_type = factor(cell_type, levels = c(cell_type_levels, "Total")),
    is_total = Sample == "Total" | cell_type == "Total",
    fill_value = ifelse(is_total, NA, Nuclei),

    # force row order manually
    y_pos = dplyr::case_when(
      Sample == "d0" ~ 3,
      Sample == "d14" ~ 2,
      Sample == "Total" ~ 1
    )
  )

n_x <- length(c(cell_type_levels, "Total"))

ggplot(plot_df, aes(x = cell_type, y = y_pos)) +
  geom_tile(
    data = dplyr::filter(plot_df, !is_total),
    aes(fill = fill_value),
    color = NA
  ) +
  geom_tile(
    data = dplyr::filter(plot_df, is_total),
    fill = "grey90",
    color = NA
  ) +
  geom_vline(
    xintercept = seq(1.5, n_x - 0.5, by = 1),
    colour = "grey75",
    linewidth = 0.8
  ) +
  geom_hline(
    yintercept = c(1.5, 2.5),
    colour = "grey75",
    linewidth = 0.8
  ) +
  geom_text(aes(label = Nuclei), size = 3.4, fontface = "bold") +
  scale_y_continuous(
    breaks = c(3, 2, 1),
    labels = c("d0", "d14", "Total"),
    limits = c(0.5, 3.5),
    expand = c(0, 0)
  ) +
  scale_fill_gradient(
    low = "grey95",
    high = "#A63A2A",
    na.value = "grey90",
    guide = "none"
  ) +
  labs(
    title = "Number of nuclei per cell type and sample",
    x = NULL,
    y = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    axis.line = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_text(face = "bold", angle = 45, hjust = 1, vjust = 1),
    axis.text.y = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.5)
  )
```

## Save final objects

```{r save RVS final, eval=FALSE}
qs_save(
  RVS,
  file = here::here(output_dir_data, "RVS.qs2"),
  nthreads = 32
)
```

```{r save RVS RNA, eval=FALSE}
RVS_RNA <- RVS
RVS_RNA@assays[["SCT"]] <- NULL
DefaultAssay(RVS_RNA) <- "RNA"

qs_save(
  RVS_RNA,
  file = here::here(output_dir_data, "RVS_RNA.qs2"),
  nthreads = 32
)

rm(RVS_RNA)
```
